---
title: "Getting Started with BasicStatsHelper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with BasicStatsHelper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(BasicStatsHelper)
library(ggplot2)
```

# Introduction

The `BasicStatsHelper` package provides simple, easy-to-use functions for common statistical analyses. This vignette demonstrates the main functions with practical examples.

## Generating Synthetic Data

Let's start by generating some synthetic data to work with:

```{r generate-data}
# Generate normal data
normal_data <- generate_synthetic_data(n = 200, type = "normal", seed = 123)
cat("Generated", length(normal_data$data), "observations from a", normal_data$distribution, "distribution\n")
cat("True mean:", normal_data$true_mean, ", True SD:", round(normal_data$true_sd, 2), "\n")

# Generate skewed data for comparison
skewed_data <- generate_synthetic_data(n = 200, type = "skewed", seed = 456)
cat("\nGenerated", length(skewed_data$data), "observations from a", skewed_data$distribution, "distribution\n")
```

## Descriptive Statistics

Calculate comprehensive descriptive statistics:

```{r descriptive-stats}
# Analyze normal data
normal_stats <- descriptive_stats(normal_data$data)
print(normal_stats)

# Analyze skewed data
skewed_stats <- descriptive_stats(skewed_data$data)
cat("\nSkewed data statistics:\n")
cat("Mean:", round(skewed_stats$mean, 2), "\n")
cat("Median:", round(skewed_stats$median, 2), "\n")
cat("Standard deviation:", round(skewed_stats$sd, 2), "\n")
```

## Confidence Intervals

Calculate confidence intervals for the population mean:

```{r confidence-intervals}
# 95% confidence interval for normal data
ci_normal <- confidence_interval(normal_data$data, conf.level = 0.95)
cat("95% Confidence Interval for Normal Data:\n")
cat("Sample mean:", round(ci_normal$mean, 2), "\n")
cat("CI: [", round(ci_normal$lower, 2), ",", round(ci_normal$upper, 2), "]\n")
cat("Margin of error:", round(ci_normal$margin_error, 2), "\n")

# Compare with different confidence levels
ci_99 <- confidence_interval(normal_data$data, conf.level = 0.99)
cat("\n99% Confidence Interval:\n")
cat("CI: [", round(ci_99$lower, 2), ",", round(ci_99$upper, 2), "]\n")
cat("Wider interval due to higher confidence level\n")
```

## Hypothesis Testing

Perform t-tests with clear interpretations:

```{r hypothesis-testing}
# One-sample t-test: Is the mean significantly different from 50?
ttest_one <- simple_ttest(normal_data$data, mu = 50)
cat("One-sample t-test results:\n")
cat(ttest_one$interpretation)

# Two-sample t-test: Compare normal vs skewed data
ttest_two <- simple_ttest(normal_data$data, skewed_data$data)
cat("\n\nTwo-sample t-test results:\n")
cat(ttest_two$interpretation)
```

## Data Visualization

Create publication-ready plots:

```{r visualization, fig.width=8, fig.height=6}
# Basic histogram
p1 <- plot_histogram(normal_data$data, 
                    title = "Normal Distribution Sample",
                    xlab = "Value",
                    bins = 25)
print(p1)

# Histogram with normal overlay and statistics
p2 <- plot_histogram(skewed_data$data,
                    title = "Skewed Distribution with Statistics",
                    add_normal = TRUE,
                    add_stats = TRUE,
                    color = "lightcoral",
                    bins = 30)
print(p2)
```

## Complete Analysis Workflow

Here's a complete example of analyzing a dataset:

```{r complete-workflow}
# Generate a new dataset
test_data <- generate_synthetic_data(n = 150, type = "bimodal", seed = 789)

# Step 1: Descriptive statistics
desc_stats <- descriptive_stats(test_data$data)
cat("Dataset Summary:\n")
cat("- Sample size:", desc_stats$n, "\n")
cat("- Mean:", round(desc_stats$mean, 2), "\n")
cat("- Median:", round(desc_stats$median, 2), "\n")
cat("- Standard deviation:", round(desc_stats$sd, 2), "\n")

# Step 2: Confidence interval
ci <- confidence_interval(test_data$data)
cat("\n95% Confidence Interval: [", round(ci$lower, 2), ",", round(ci$upper, 2), "]\n")

# Step 3: Hypothesis test
ttest <- simple_ttest(test_data$data, mu = 45)
cat("\nHypothesis test (H0: μ = 45):\n")
cat("P-value:", round(ttest$p.value, 4), "\n")
cat("Significant?", ttest$significant, "\n")

# Step 4: Visualization
plot_histogram(test_data$data,
              title = "Bimodal Distribution Analysis",
              add_stats = TRUE,
              color = "lightgreen")
```

## Conclusion

The `BasicStatsHelper` package provides a streamlined workflow for common statistical analyses:

1. **`generate_synthetic_data()`** - Create test datasets
2. **`descriptive_stats()`** - Calculate comprehensive summary statistics
3. **`confidence_interval()`** - Estimate population parameters
4. **`simple_ttest()`** - Perform hypothesis testing with clear interpretations
5. **`plot_histogram()`** - Create publication-ready visualizations

These functions are designed to be intuitive for beginners while providing the statistical rigor needed for research applications.
